---
title: "R Notebook"
output: html_document
author: "Dhruv Dole, Ahmed Gharib"
---
# Is there a relationship between students' socioeconomic status and there academic test scores?
```{r}
library(tidyverse)
raw_data <- read.csv("https://raw.githubusercontent.com/dsdole/DS202Final/main/data/Expanded_data_with_more_features.csv")
```
# Erasing "X" from the dataset:
```{r}
raw_data <- select(raw_data, -X)
head(raw_data)
```
```{r}
raw_data$Gender <- as.factor(raw_data$Gender)
raw_data$EthnicGroup <- as.factor(raw_data$EthnicGroup)
raw_data$ParentEduc <- as.factor(raw_data$EthnicGroup)
raw_data$LunchType <- as.factor(raw_data$LunchType)
raw_data$TestPrep <- as.factor(raw_data$TestPrep)
raw_data$ParentMaritalStatus <- as.factor(raw_data$ParentMaritalStatus)
raw_data$TransportMeans <- as.factor(raw_data$TransportMeans)

#raw_data$isFirstChild <- as.factor(raw_data$isFirstChild)
```

# Factoring WklyStudyHours
```{r}
unique(raw_data$WklyStudyHours)
raw_data$WklyStudyHours <- factor(raw_data$WklyStudyHours, ordered = TRUE,
                                  levels = c("< 5", "5 - 10", "> 10"))
```

# Filling empty space
```{r}
cleanraw_data <- raw_data %>%
  mutate_all(~replace(., . == "", NA))
cleanraw_data %>%
  summarise_all(~mean(is.na(.)) * 100)
```


# Visual of clean data
```{r}
head(cleanraw_data)
```

# Correlation plots (Ahmed Gharib)
## 1 - Average score affected by students according to their Practicing sports status:
```{r}
library(ggplot2)

AvgScore <- rowMeans(cleanraw_data[, c("MathScore", "ReadingScore", "WritingScore")], na.rm = TRUE)


ggplot(cleanraw_data, aes(x = PracticeSport, y = AvgScore, fill = Gender)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge", alpha = 0.8) +
  scale_fill_manual(values = c("orange", "purple")) +
  labs(x = "Practicing sports status", y = "Average Score", fill = "Gender") +
  theme_bw()

```
As we can see above the Average overall scores affected by students who practice sports compared to those who do not, is not a huge difference. Very similar overall.

## 2 - Weekly study hours affection of Average scores

```{r}
# Load required packages
# library(ggplot2)

# Compute average score
AvgScore <- rowMeans(cleanraw_data[, c("MathScore", "ReadingScore", "WritingScore")], na.rm = TRUE)

ggplot(cleanraw_data, aes(x = AvgScore, fill = factor(WklyStudyHours))) +
  geom_histogram(binwidth = 2, position = "dodge") +
  scale_fill_manual(values = c("black", "blue", "red")) +
  facet_wrap(~WklyStudyHours) + # remove this line to remove facets
  labs(title = "Distribution of Average Scores by Weekly Study Hours",
       x = "Average Score",
       y = "Student Count",
       fill = "Weekly Study Hours") +
  theme_minimal()

```
For the above histogram, gathering the distribution of weekly hours of study seeing how that affects the overall scores of the students. As seen students that study less than 5 hours a week, have a peak range around 60-75. Similarly, students who study for more than 10 hours, have around the same range peak. But less students. Seems like 5-10 hours of study weekly is the best amount of hours in order to get better scores for the more students in this data set.

# 3 - Correlating Number of siblings to each individual scores (Lineplot)
```{r}
# library(tidyr)

# Create a new data frame with columns for NbrSiblings, MathScore, ReadingScore, and WritingScore
scores_by_siblings <- cleanraw_data %>%
  select(NrSiblings, MathScore, ReadingScore, WritingScore) %>%
  na.omit()

# Reshape the data to long format
scores_by_siblings_long <- scores_by_siblings %>%
  pivot_longer(cols = c("MathScore", "ReadingScore", "WritingScore"), names_to = "Subject", values_to = "Score")

# Create a line plot with facets for each subject
ggplot(scores_by_siblings_long, aes(x = NrSiblings, y = Score, group = Subject)) +
  geom_line(aes(color = Subject)) +
  labs(title = "Correlation between Number of Siblings and Test Scores",
       x = "Number of Siblings",
       y = "Test Score",
       color = "Subject") +
  facet_wrap(~Subject, scales = "free_y") +
  theme_minimal()
```
```{r}

```
# Mosiac Plot between Ethnic Group and Parent Marital Status
```{r}

library(ggmosaic)
mosaic_examp <- ggplot(data = cleanraw_data) +
  geom_mosaic(aes(x = product(EthnicGroup, ParentMaritalStatus), fill = EthnicGroup)) +
  labs(y = "Ethnic Group", x = "Parent Marital Status", title = "Ethnic Group and Parent Marital Status")
mosaic_examp
```
There is no relation between Ethnic Group and Parent Marital Status. This is because there is no change in the proportions from column to column or row to row. This means that we can use these as independent predictors for model building.
```{r}
ggplot(cleanraw_data, aes(x = MathScore, y = ReadingScore, color = EthnicGroup, shape = Gender)) +
  geom_point(size = 3) +
  facet_grid(Gender ~ EthnicGroup) +
  labs(x = "Math Score", y = "Reading Score", color = "Ethnic Group", shape = "Gender") +
  theme_bw()
```
```{r}
ggplot(cleanraw_data, aes(x = TestPrep, y = MathScore, fill = Gender)) +
  geom_boxplot() +
  scale_fill_manual(values = c("blue", "green", "red")) +
  labs(x = "Group", y = "Value", fill = "Variable") +
  ggtitle("Score Ranges of males and females with varying test prep completion")
```
